---
title: "Data Pre-Processing"
author: "Goh Jie Xin"
date: "10/4/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Libraries

```{r}
library(dplyr)
library(DataExplorer)
library(mice)
library(VIM)
library(missForest)
library(caret)
```


# Read Data

```{r}
# Convert "NA" or Blanks to NA
# Convert String Type to Factor
df <- read.csv('aug_train.csv', header = T, na.strings=c("NA","NaN", ""), stringsAsFactors=T)

head(df,5)

# If the df has blanks, fill in with NA
df <- mutate_all(df, na_if, "")
```


# Remove unrelated column

1. Column enrollee_id does not give any useful information, so we can delete the column.
2. Column city and city_development_index is correlated. Same city will have same city_development_index, so we can remove column city (as it has too many categories).
```{r}
# Aggregate function (Group by and Count Unique)
plot(aggregate( df$city_development_index~df$city,df, function(x) length(unique(x))))

# Remove columns
df$enrollee_id <- NULL
df$city <- NULL
```


# Convert Column Format

1. Column target is in num type. We need to convert to factor type.
```{r}
# Convert column format to factor
df$target <- as.factor(df$target)
```

2. Column experience is factor type. We can convert to num type but before convert need to replace:
   "<1" --> "0" --> 0
   ">20" --> "21"  --> 21
```{r}
print("Before conversion: ")
levels(df$experience)
levels(df$experience) <- c(levels(df$experience),"0","21")
df$experience[df$experience=="<1"]  <- "0" 
df$experience[df$experience==">20"]  <- "21"
# Convert from factor to numeric
df$experience <- as.numeric(as.character(df$experience))
print("After conversion: ")
hist(df$experience)
```

3. Column last_new_job is factor type. We can convert to num type but before convert need to replace:
   ">4" --> "5"
   "never" --> "0"
```{r}
print("Before conversion: ")
levels(df$last_new_job)
levels(df$last_new_job) <- c(levels(df$last_new_job),"5","0")
df$last_new_job[df$last_new_job==">4"]  <- "5" 
df$last_new_job[df$last_new_job=="never"]  <- "0"
# COnvert from factor to numeric
df$last_new_job <- as.numeric(as.character(df$last_new_job))
print("After conversion: ")
hist(df$last_new_job)

write.csv(df, file = "aug_train2.csv")
```


# Handling Missing Values

Check if there is any missing values.
```{r}
# View missing values
colSums(is.na(df))
plot_missing(df)
```
There are 8 columns with missing values. Instead of removing those columns, we will try to fill in the missing values with different method.

## Method 1 - Imputation
1. Continuous column - fill NA by mean.
   na.rm = True : ignore NA when computing mean
2. Factor column - fill NA by mode
```{r}
# Method 1 - Imputation

# Impute continuous var (mean)
dfImputation <- df
contCol <- names(dfImputation)[sapply(dfImputation, is.numeric)]
for (col in contCol){
   sumNA <- sum(is.na(df[[col]]))
   dfImputation[ ,col][is.na(dfImputation[ ,col])] <- mean(dfImputation[ ,col], na.rm=TRUE)
   # Plot Density Plot Before & After Imputation
   plot(density(na.omit(df[[col]])),
        main=sprintf("Density plot of %s (Initial NA count: %d)", col,  sumNA))
   lines(density(dfImputation[[col]]),col="red")
   legend("topright", cex=0.8, horiz=TRUE,
          c("Before Imputation", "After Imputation"),
          lty = 1, lwd = 2, col = c("black", "red"))

}

# Impute factor var 
facCol <- names(dfImputation)[sapply(dfImputation, is.factor)]
for (col in facCol){
   sumNA <- sum(is.na(df[[col]]))
   mode <- function(x){levels(x)[which.max(tabulate(x))]}  # Function to obtain mode
   dfImputation[ ,col][is.na(dfImputation[ ,col])] <-  mode(df[[col]])
   # Prepare data to plot graph
   colType <- c(rep("Before Imputation", dim(df)[1]), 
                rep("After Imputation", dim(dfImputation)[1]))
   colVal <- c(as.character(df[,col]), as.character(dfImputation[,col]))
   outcome_data <- as.data.frame(unclass(cbind(colType, colVal)),
                                 stringsAsFactors = TRUE)
   outcome_data <- data.frame(Type = factor(colType, 
                                            levels = c("Before Imputation", "After Imputation")),
                                            Category = colVal)
   # Plot side by side bar chart
   print(ggplot(data = data.frame(table(outcome_data)), aes(x = Category, y = Freq, fill = Type)) +
          geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
          geom_text(aes(label = Freq), fontface = "bold", vjust = 1.5,
            position = position_dodge(.9), size = 4) +
          labs(x = col, y = "Frequency", 
               title = sprintf(" %s (Initial NA count: %d)", col,  sumNA)) +
         theme(plot.title = element_text(hjust = 0.5),
               axis.title.x = element_text(face="bold", colour="black", size = 12),
               axis.title.y = element_text(face="bold", colour="black", size = 12),
               legend.title = element_text(face="bold", size = 10)))

}

```


## Method 2 - MICE
Setting: 5 iterations, cart (imputation by classification & regression trees)
As MICE method takes time to run, the output is exported. (aug_train_mice.csv)
```{r}
# Method 2 - MICE
# mice return list, use complete() to convert to df
# df_impute_mice <- mice(df, 5, method="cart")
# df_impute_mice <- complete(df_impute_mice)
# plot_missing(df_impute_mice)
# write.csv(df_impute_mice, file = "aug_train_mice.csv")
df_impute_mice <- read.csv('aug_train_mice.csv', header = T, na.strings=c("NA","NaN", ""), stringsAsFactors=T)
```


## Method 3 - missForest
Setting: 5 iterations
As missForest method takes time to run, the output is exported (aug_train_mF.csv)
```{r}
# Method 3 - missForest
# df_mF <- missForest(df, maxiter = 5)
# df_mF <- (df_mF$ximp)
# plot_missing(df_mF)
# write.csv(df_mF, file = "aug_train_mF.csv")
df_mF <- read.csv('aug_train_mF.csv', header = T, na.strings=c("NA","NaN", ""), stringsAsFactors=T)
```

Compare 3 methods of imputation.
```{r}
# Plot graph

# Continuous Column
contCol <- names(dfImputation)[sapply(dfImputation, is.numeric)]
for (col in contCol){
   sumNA <- sum(is.na(df[[col]]))
   # Plot Density Plot Before & After Imputation
   plot(density(na.omit(df[[col]])),
        main=sprintf("Density plot of %s (Initial NA count: %d)", col,  sumNA))
   lines(density(dfImputation[[col]]),col="red")
   lines(density(df_impute_mice[[col]]),col="blue")
   lines(density(df_mF[[col]]),col="brown")
   legend("topright", cex=0.7, horiz=TRUE,
          c("Original", "Impute by Mean", "MICE", "missForest"),
          lty = 1, lwd = 2, col = c("black", "red", "blue", "green"))
}

# Categorical variable
# Impute factor var 
facCol <- names(dfImputation)[sapply(dfImputation, is.factor)]
for (col in facCol){
   sumNA <- sum(is.na(df[[col]]))
   # Prepare data to plot graph
   colType <- c(rep("Original", dim(df)[1]), 
                rep("Impute by Mode", dim(dfImputation)[1]),
                rep("MICE", dim(df_impute_mice)[1]),
                rep("missForest", dim(df_mF)[1]))
   colVal <- c(as.character(df[,col]), as.character(dfImputation[,col]),
               as.character(df_impute_mice[,col]), as.character(df_mF[,col]))
   outcome_data <- as.data.frame(unclass(cbind(colType, colVal)),
                                 stringsAsFactors = TRUE)
   outcome_data <- data.frame(Type = factor(colType, 
                                            levels = c("Original", "Impute by Mode", 
                                                       "MICE", "missForest")),
                                            Category = colVal)
   # Plot side by side bar chart
   print(ggplot(data = data.frame(table(outcome_data)), aes(x = Category, y = Freq, fill = Type)) +
          geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
          geom_text(aes(label = Freq), vjust = 1.5,
            position = position_dodge(.9), size = 2) +
          labs(x = col, y = "Frequency", 
               title = sprintf(" %s (Initial NA count: %d)", col,  sumNA)) +
         theme(plot.title = element_text(hjust = 0.5),
               axis.title.x = element_text(face="bold", colour="black", size = 12),
               axis.title.y = element_text(face="bold", colour="black", size = 12),
               legend.title = element_text(face="bold", size = 10)))
}
```


**The following will use df_impute_mice instead of the original df.


# Data Encoding
Reference: https://www.r-bloggers.com/2020/02/a-guide-to-encoding-categorical-features-using-r/
Encoding for categorical features:
Nominal - One Hot Encoding
Ordinal - defined order

## Encoding - Ordinal Variables
Ordinal variables: education_level, company_size
```{r}
ordinalCol = c("education_level", "company_size")

# Encoding for Ordinal Variable, Follow defined order
encode_ordinal <- function(x, order = unique(x)) {
  x <- as.numeric(factor(x, levels = order, exclude = NULL))
  x
}

df_impute_mice_encoded <- df_impute_mice

# Column education_level
levels(df_impute_mice_encoded$education_level)
table(df_impute_mice_encoded[["education_level"]],
      encode_ordinal(df_impute_mice_encoded[["education_level"]],
      order = c("Primary School", "High School", "Graduate", "Masters", "Phd")),
      useNA = "ifany")

# Column company_size
levels(df_impute_mice_encoded$company_size)
table(df_impute_mice_encoded[["company_size"]],
      encode_ordinal(df_impute_mice_encoded[["company_size"]],
      order = c("<10", "10/49", "50-99", "100-500", "500-999", 
                "1000-4999", "5000-9999", "10000+")),
      useNA = "ifany")

# Encode the df
df_impute_mice_encoded[["education_level"]] <- encode_ordinal(df_impute_mice_encoded[["education_level"]],
               order = c("Primary School", "High School", "Graduate", "Masters", "Phd"))
df_impute_mice_encoded[["company_size"]] <- encode_ordinal(df_impute_mice_encoded[["company_size"]],
               order = c("<10", "10/49", "50-99", "100-500", "500-999", 
                         "1000-4999", "5000-9999", "10000+"))

for (col in ordinalCol){
   hist((df_impute_mice_encoded[[col]]), main=col, xlab=col)
}

```


## Encoding - Nominal Variables
Nominal variables: gender, enrolled_university, major_discipline, company_type
Use dummyVars from caret Package
```{r}
nominalCol = c("gender", "enrolled_university", "major_discipline", "company_type")
dummy <- caret::dummyVars(" ~ gender + enrolled_university + major_discipline + company_type", 
                          data = df_impute_mice_encoded)
dummy <- data.frame(predict(dummy, newdata = df_impute_mice_encoded))
df_impute_mice_encoded <- cbind(subset(df_impute_mice_encoded, select=-c(gender, enrolled_university, major_discipline, company_type)), dummy)
head(df_impute_mice_encoded, 5)
```
Other methods:
1. mutate() by own dict
2. factor(level, label)
3. onehotencoding - dummify

## Encoding - Binary
Binary category variable: relevent_experience
"Has relevent experience" --> 1
"No relevent experience" --> 0
```{r}
levels(df_impute_mice_encoded$relevent_experience) <- c(levels(df_impute_mice_encoded$relevent_experience),"1","0")
df_impute_mice_encoded$relevent_experience[df_impute_mice_encoded$relevent_experience=="Has relevent experience"]  <- "1" 
df_impute_mice_encoded$relevent_experience[df_impute_mice_encoded$relevent_experience=="No relevent experience"]  <- "0"
# Convert from factor to numeric
df_impute_mice_encoded$relevent_experience <- as.numeric(as.character(df_impute_mice_encoded$relevent_experience))
hist(df_impute_mice_encoded$relevent_experience)
```

After encoding: 
```{r}
head(df_impute_mice_encoded, 5)
write.csv(df_impute_mice_encoded, file = "aug_train_mice_encoded.csv")
```


# Normalization
min-max normalization
Columns to be normalized: city_development_index, training_hours, experience
```{r}
normCol = c("city_development_index", "training_hours", "experience")

#define Min-Max normalization function
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}

df_impute_mice_encoded_norm <- df_impute_mice_encoded
df_impute_mice_encoded_norm[normCol] <- as.data.frame(lapply(df_impute_mice_encoded[normCol], min_max_norm))

summary(df_impute_mice_encoded_norm[normCol])

for (col in normCol){
   plot(density(df_impute_mice_encoded_norm[[col]]), main=col)
}
```


Finally, export the clean data (aug_train_clean.csv)
```{r}
write.csv(df_impute_mice_encoded_norm, file = "aug_train_clean.csv")
```
